Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface

### 2022-04-05 doc/AMS-15/AMS-15_paper/AMS_HGIS_els/AMS_HGIS_els.Rmd
add elsevier rticles format

---
title: "Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface"
author:
- Brett Longworth
- Josh Burton
- Simon Pendleton
- Sydney Moser
- Mark Roberts
- Mark Kurz
output:
  pdf_document: default
  word_document: default
  html_document: default
bibliography: HGIS_paper.bib
csl: elsevier-without-titles.csl
---

@@

# Abstract

We have developed a method for measuring radiocarbon in carbonate minerals as CO~2~ gas via a NEC MCSNICS hybrid sputter gas ion source (HGIS). The method uses helium as a carrier gas to displace CO~2~ from sample vials to an open split, where a glass capillary samples the mixture for delivery directly to the HGIS. This method skips the gas transfer and quantification steps used in a closed inlet HGIS system, simplifying sample measurement. Samples larger than 8 mg carbonate can be measured. Results from measurements of consensus standards (TIRI-I, IAEA C2, and an internal modern shell standard), and samples from a marine core (F^14^C = 0.4 - 1.15) show that the method agrees well with traditional AMS measurement of the same samples as graphite, and that the 1σ uncertainty is about 1 %. We discuss advantages and disadvantages of continuous flow sample introduction, and the effect of reduced precision on calibrated age-depth models produced using gas-source data.

@@

The open split interface (Figure 1) supplies a helium - CO~2~ mixture to a split where a capillary delivers gas to the source and the rest is vented. Flowing helium is used to displace sample CO~2~ from the sample vial to the delivery capillary in the split. Our open-split interface is constructed of simple, off-the-shelf parts. The core of the design is a dual needle and split housed in a VICI cross fitting. Displacement helium flows in through one side of the cross and into the sample vial through 1/16 in. stainless steel tubing, which forms the outer flow path of the concentric dual needle.

@@

The helium displacement flow rate was set to 250 μl min^-1^ for initial tests, and this was reduced to 100 μl min^-1^ for later work, including most reference standard comparisons and the core comparison. The flow rate from the open split into the source was defined by 150 cm of 50 μm fused silica capillary, split into two sections by a zero-volume bulkhead connector at the source high-voltage cage. The section inside the cage was jacketed in 1/16 in. PFTE tubing to help protect against spark damage. The total flow rate of gas to the source varied due to the viscosity difference between helium and CO~2~ and helium, and was estimated to be between 30 μl min^-1^ for pure helium and 40 μl min^-1^ for pure CO~2~ using the Hagen-Poiseuille equation.

@@

curmean <- mean(currents$c12ua) %>% format(digits = 2) %>% paste("μA")
cursd <- sd(currents$c12ua)
```

Initial work with the NEC source established optimal source parameters for consistent source performance. We used tank CO~2~ connected through a regulator and glass capillary to the ion source to measure source performance over a range of CO~2~ flows. Best currents for the source was 10 μA ^12^C^-^, or roughly 1/10 of graphite target current. Peak current was produced at a CO~2~ flow rate of 7.5 μl min^-1^ CO~2~, but the source produced near optimal currents with flows from 7.5-25 μl min^-1^ CO~2~. This is comparable with other hybrid sources [@Xu2007], although improved target design and continued optimization of source parameters may improve performance [@Salazar2013; @Fahrni2013]. Typical currents for this study were between 4-9 μA and average current for all samples was `r curmean` ± `r cursd`. Several studies have shown best currents at lower CO~2~ flow rates [@Xu2007; @Ruff2010].

@@

A contrast of the open split approach with closed-source HGIS methods is changing CO~2~ flow and helium dilution ratio over the period of measurement due to the dilution of the sample by the displacement helium. We examined the effect of this variable carbon delivery and dilution by measuring ^12^C^-^ current evolution during helium displacement/dilution of vial CO~2~ (Figure 2a) and comparing this to the composition of the gas delivered to the source. We estimated gas composition (Figure 2b) using a simple mixing model:

@@

The HGIS method produces higher ratios for radiocarbon dead samples than the graphite method. Measurements of radiocarbon-free reference materials (IAEA C1 marble and TIRI F spar calcite) show that the average normalized ratio for these samples is `r blankmean` ± `r blanksd` (1σ uncertainty, N=`r nblank`). We investigated the sources of this method blank as a current-dependent blank and as a current-independent blank, and implemented a blank correction for the method.

HGIS cathodes produce measurable ion current, even with no gas flow. This current is initially \> 10 μA and drops exponentially, reaching a baseline of ~100 nA after 5-10 minutes. For the range of currents in this study (3-9 μA), this baseline current represents a 1-6% contribution to the total measured sample current at the F^14^C of the target current baseline. This can be interpreted as a constant current dependent blank, analogous to a mass-dependent "constant contaminant" blank and may be modeled with the same tools [@haghipour_compound-specific_2019; @hanke_comprehensive_2017; @roberts_14c_2019; @sun_14_2020]. To assess blank contribution and ratio dependence on sample current, vials of live and dead standard gas were measured as the displacement helium flow diluted the samples, producing diminishing current over time. We used these data to develop a current-balance model following methods used to assess mass-dependent or constant contamination blanks in graphite targets [@shah_walter_ultra-small_2015; @sun_14_2020]. The constant contamination blank was also assessed directly by measuring pure helium in vials using the same method. Results from both tests are shown in Figure 3a. The inverse of the sample current is used to linearize the data, allowing assessment with a simple linear model. Fits from modern and dead dilution tests should trend towards the fraction modern of the blank, crossing at the inverse of blank current. Based on this assessment, we estimate the ratio of the blank as having F^14^C of 0.43 with a current of 110 nA.

@@

# Summary table
# cons_df %>% 
#   group_by(Name, fm_consensus) %>% 
#   summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
#                    list(mean = mean, sd = sd)),
#             N = n()) %>% 
#   ungroup() %>% 
#   arrange(fm_consensus) %>% 
#   mutate (fm_corr_err = set_errors(fm_corr_mean, fm_corr_sd),
#           Name = case_when(Name == "TIRI-I" ~ "TIRI I",
#                            Name == "C-2" ~ "IAEA C2",
#                            Name == "NOSAMS2" ~ "NOSAMS 2")) %>% 
#   select(`Sample Name` = Name, 
#          `Reference F^14^C` = fm_consensus,
#          `HGIS F^14^C ± stdev` = fm_corr_err,
#          `Mean per-sample error` = sig_fm_corr_mean,
#          N) %>% 
#   knitr::kable(digits = 3,
#                caption = "Reference material precision and accuracy. Reference 
#                values are community consensus values for TIRI I and C2, and the 
#                NOSAMS internal graphite value for NOSAMS2. NOSAMS2 is a 
#                homogenized bivalve molusk reference. The mean of per-sample 
#                error estimates is included for comparison to the reproducibilty 
#                of the samples.")

  # gt() %>% 
  # tab_header(title = "Reference material precision and accuracy",
  #            subtitle = "Values in pMC, Errors are SD of measurements") %>% 
  # fmt_number(4, 
  #            decimals = 3) %>% 
  # cols_align("right", 3)

@@

Reference materials run with the hybrid gas method were in good agreement with consensus values. Mean difference from consensus for all measurements was F^14^C `r format(cons_sum$fm_diff_mean, digits = 1, nsmall = 4)` shows that the hybrid method has no bias in aggregate. The standard deviation of the difference in F^14^C was `r format(cons_sum$fm_diff_sd, nsmall = 4)`, which may be taken as an estimate of the 1σ precision of the method. The mean of propagated measurement errors for these samples was `r format(cons_sum$sig_fm_corr_mean, nsmall = 4)` F^14^C for these samples, which is in good agreement with the method precision above. The method error is about three times larger than the roughly 3 ‰ precision of graphite measurements, but is within the range of method precision of other gas sources.

The difference between reference samples and graphite consensus does not vary systematically by F^14^C, indicating that the method has no significant bias with radiocarbon content (Figure 4). Each reference standard type agrees with graphite consensus values within the standard deviation of samples measured, with slightly more measurements agreeing with graphite within 1σ uncertainties than would be expected statistically.

@@

We found the HGIS and graphite measurements to be in good agreement, with no significant bias over the range in F^14^C of the core samples (0.4 to 1.15) (Figure 5a). Mean difference between hybrid and graphite measurements was -0.003 F^14^C and the standard deviation of this difference was 0.009 F^14^C. Mean per-sample 1σ uncertainty was 0.010 F^14^C for HGIS measurements, compared with 0.002 F^14^C for graphite. Three samples were analysed by HGIS in duplicate. Two of the three pairs agreed within 1σ and all three agreed within 2σ. There is an apparent, but not significant positive relationship between the HGIS-graphite difference and measured value. The higher graphite values at lower F^14^C may indicate an overcorrection of HGIS blank for these samples.

@@

# Tables

```{r message = FALSE}
cons_df %>% 
  group_by(Name, fm_consensus) %>% 
  summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  ungroup() %>% 
  arrange(fm_consensus) %>% 
  mutate (fm_corr_err = set_errors(fm_corr_mean, fm_corr_sd),
          Name = case_when(Name == "TIRI-I" ~ "TIRI I",
                           Name == "C-2" ~ "IAEA C2",
                           Name == "NOSAMS2" ~ "NOSAMS 2")) %>% 
  select(`Sample Name` = Name, 
         `Reference F^14^C` = fm_consensus,
         `HGIS F^14^C ± stdev` = fm_corr_err,
         `Mean per-sample error` = sig_fm_corr_mean,
         N) %>% 
  knitr::kable(digits = 3,
               caption = "Table 1. Reference material precision and accuracy. Reference 
               values are community consensus values for TIRI I and C2, and the 
               NOSAMS internal graphite value for NOSAMS2. NOSAMS2 is a 
               homogenized bivalve molusk reference. The mean of per-sample 
               error estimates is included for comparison to the reproducibilty 
               of the samples.")

# Figures

![Figure 1](images/HGIS_open_split_drawing.svg)

![Figure 2](figures/fig2_current_flow.svg)

![Figure 3](figures/fig3_blanks.svg)

![Figure 4](figures/fig4_consensus_samples.svg)

![Figure 5](figures/fig5_core_carbonates.svg)

# Figure captions

Figure 1. Schematic of the open split inlet. Two valco valves provide the ability to admit helium and live or dead tank standards. The double needle is housed in a VICI cross and allows displacement helium to flow into the sample vial and the sample-helium mixture to flow to the open split, where a capillary delivers the mixture to the ion source.

Figure 2. Performance of the NOSAMS MCSNICS hybrid gas ion source. (a) Measured ^12^C^-^ ion current during displacement of a pure CO~2~ vial sample with helium. (b) Modeled flows of helium and CO~2~ to the ion source during the same dilution experiment. (c) Measured ion current for modeled CO~2~ flow. The stability of the source over a range of flow produces stable current for up to 40 min of measurement time.

Figure 3. Assessment of blanks. (a) Current dependent blank model and data. F^14^C and current of measurements from dilution tests of live and dead gas are used to generate linear fits which should cross at the F^14^C and current of the constant current blank. Direct determination of F^14^C and current of the blank are also shown. (b) Fraction modern of large, radiocarbon dead samples compared to sample current. Horizontal and vertical bars represent measurement uncertainty in current and ratio, respectively.

Figure 4. Agreement of standards with graphite consensus values. Difference of HGIS measurements from consensus for individual samples are plotted against their expected value. Expected values are spread to avoid overlapping points. Error bars represent one sigma sample uncertainty. The black line is the mean difference from expected value for all measurements, and the blue lines are one standard deviation of all measurements.

Figure 5. Comparison of Haiti core samples measured as graphite and via HGIS. (a) Difference between HGIS and graphite measurements of the same materials. Difference between HGIS and graphite measurements is shown on the vertical axis and plotted against the mean of the two measurements. Samples are offset on the horizontal axis for clarity. Error bars are measurement error for each sample. Graphite measurements are included to allow comparison of error estimates. The bottom row of plots show a comparison of age-depth models produced using HGIS (b) and graphite measurements (c).

# References

>>>>

---
title: "Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface"
author:
  - name: Brett Longworth
    email: blongworth@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
    correspondingauthor: true
  - name: Josh Burton
    email: jburton@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
  - name: Simon Pendleton
    email: simon.pendleton@plymouth.edu
    affiliation: Plymouth State University
  - name: Sydney Moser
    email: smose007@uottawa.ca
    affiliation: University of Ottawa
  - name: Mark Roberts
    email: mroberts@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
  - name: Mark Kurz
    email: mkurz@whoi.edu
    affiliation: Woods Hole Oceanographic Institution
address:
  - code: Woods Hole Oceanographic Institution
    address: Department of Geology and Geophysics, 360 Woods Hole Rd., Woods Hole, MA, 02540
  - code: Plymouth State University
    address: Department, Street, City, State, Zip
  - code: University of Ottawa
    address: Department, Street, City, State, Zip
footnote:
abstract: |
  We have developed a method for measuring radiocarbon in carbonate minerals as CO~2~ gas via a NEC MCSNICS hybrid sputter gas ion source (HGIS). The method uses helium as a carrier gas to displace CO~2~ from sample vials to an open split, where a glass capillary samples the mixture for delivery directly to the HGIS. This method skips the gas transfer and quantification steps used in a closed inlet HGIS system, simplifying sample measurement. Samples larger than 8 mg carbonate can be measured. Results from measurements of consensus standards (TIRI-I, IAEA C2, and an internal modern shell standard), and samples from a marine core (F^14^C = 0.4 - 1.15) show that the method agrees well with traditional AMS measurement of the same samples as graphite, and that the 1$\sigma$ uncertainty is about 1 %. We discuss advantages and disadvantages of continuous flow sample introduction, and the effect of reduced precision on calibrated age-depth models produced using gas-source data.
keywords: 
  - radiocarbon
  - carbonate
  - hybrid gas ion source
journal: "Nuclear Instruments and Methods in Physics Research Section B"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: HGIS_paper.bib
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard
csl: elsevier-without-titles.csl
output: 
  rticles::elsevier_article:
    keep_tex: true
    citation_package: default
---

@@

@@

The open split interface (Figure \ref{fig1}) supplies a helium - CO~2~ mixture to a split where a capillary delivers gas to the source and the rest is vented. Flowing helium is used to displace sample CO~2~ from the sample vial to the delivery capillary in the split. Our open-split interface is constructed of simple, off-the-shelf parts. The core of the design is a dual needle and split housed in a VICI cross fitting. Displacement helium flows in through one side of the cross and into the sample vial through 1/16 in. stainless steel tubing, which forms the outer flow path of the concentric dual needle.

@@

The helium displacement flow rate was set to 250 $\mu$l min^-1^ for initial tests, and this was reduced to 100 $\mu$l min^-1^ for later work, including most reference standard comparisons and the core comparison. The flow rate from the open split into the source was defined by 150 cm of 50 $\mu$m fused silica capillary, split into two sections by a zero-volume bulkhead connector at the source high-voltage cage. The section inside the cage was jacketed in 1/16 in. PFTE tubing to help protect against spark damage. The total flow rate of gas to the source varied due to the viscosity difference between helium and CO~2~ and helium, and was estimated to be between 30 $\mu$l min^-1^ for pure helium and 40 $\mu$l min^-1^ for pure CO~2~ using the Hagen-Poiseuille equation.

```{r fig1, fig.width = 5, fig.height = 5, fig.align='center', out.width="50%", fig.cap = "\\label{fig1}Schematic of the open split inlet. Two valco valves provide the ability to admit helium and live or dead tank standards. The double needle is housed in a VICI cross and allows displacement helium to flow into the sample vial and the sample-helium mixture to flow to the open split, where a capillary delivers the mixture to the ion source.", echo = FALSE}
knitr::include_graphics("../figures/fig1_HGIS_open_split_drawing.pdf")
```

@@

curmean <- mean(currents$c12ua) %>% format(digits = 2)
cursd <- sd(currents$c12ua)
```

Initial work with the NEC source established optimal source parameters for consistent source performance. We used tank CO~2~ connected through a regulator and glass capillary to the ion source to measure source performance over a range of CO~2~ flows. Best currents for the source was 10 $\mu$A ^12^C^-^, or roughly 1/10 of graphite target current. Peak current was produced at a CO~2~ flow rate of 7.5 $\mu$l min^-1^ CO~2~, but the source produced near optimal currents with flows from 7.5-25 $\mu$l min^-1^ CO~2~. This is comparable with other hybrid sources [@Xu2007], although improved target design and continued optimization of source parameters may improve performance [@Salazar2013; @Fahrni2013]. Typical currents for this study were between 4-9 $\mu$A and average current for all samples was `r curmean` $\mu$A ± `r cursd`. Several studies have shown best currents at lower CO~2~ flow rates [@Xu2007; @Ruff2010].

```{r fig2, fig.width = 5, fig.height = 5, fig.align='center', out.width="50%", fig.cap = "\\label{fig2}Performance of the NOSAMS MCSNICS hybrid gas ion source. (a) Measured $^{12}C^-$ ion current during displacement of a pure $CO_2$ vial sample with helium. (b) Modeled flows of helium and $CO_2$ to the ion source during the same dilution experiment. (c) Measured ion current for modeled $CO_2$ flow. The stability of the source over a range of flow produces stable current for up to 40 min of measurement time.", echo = FALSE}
knitr::include_graphics("../figures/fig2_current_flow.pdf")
```

@@

A contrast of the open split approach with closed-source HGIS methods is changing CO~2~ flow and helium dilution ratio over the period of measurement due to the dilution of the sample by the displacement helium. We examined the effect of this variable carbon delivery and dilution by measuring ^12^C^-^ current evolution during helium displacement/dilution of vial CO~2~ (Figure \ref{fig2}a) and comparing this to the composition of the gas delivered to the source. We estimated gas composition (Figure \ref{fig2}b) using a simple mixing model:

@@

The HGIS method produces higher ratios for radiocarbon dead samples than the graphite method. Measurements of radiocarbon-free reference materials (IAEA C1 marble and TIRI F spar calcite) show that the average normalized ratio for these samples is `r blankmean` ± `r blanksd` (1$\sigma$ uncertainty, N=`r nblank`). We investigated the sources of this method blank as a current-dependent blank and as a current-independent blank, and implemented a blank correction for the method.

HGIS cathodes produce measurable ion current, even with no gas flow. This current is initially \> 10 $\mu$A and drops exponentially, reaching a baseline of ~100 nA after 5-10 minutes. For the range of currents in this study (3-9 $\mu$A), this baseline current represents a 1-6% contribution to the total measured sample current at the F^14^C of the target current baseline. This can be interpreted as a constant current dependent blank, analogous to a mass-dependent "constant contaminant" blank and may be modeled with the same tools [@haghipour_compound-specific_2019; @hanke_comprehensive_2017; @roberts_14c_2019; @sun_14_2020]. To assess blank contribution and ratio dependence on sample current, vials of live and dead standard gas were measured as the displacement helium flow diluted the samples, producing diminishing current over time. We used these data to develop a current-balance model following methods used to assess mass-dependent or constant contamination blanks in graphite targets [@shah_walter_ultra-small_2015; @sun_14_2020]. The constant contamination blank was also assessed directly by measuring pure helium in vials using the same method. Results from both tests are shown in Figure 3a. The inverse of the sample current is used to linearize the data, allowing assessment with a simple linear model. Fits from modern and dead dilution tests should trend towards the fraction modern of the blank, crossing at the inverse of blank current. Based on this assessment, we estimate the ratio of the blank as having F^14^C of 0.43 with a current of 110 nA.

```{r fig3, fig.width = 5, fig.height = 5, fig.align='center', out.width="50%", fig.cap = "\\label{fig3}Assessment of blanks. (a) Current dependent blank model and data. $F^{14}C$ and current of measurements from dilution tests of live and dead gas are used to generate linear fits which should cross at the $F^{14}C$ and current of the constant current blank. Direct determination of $F^{14}C$ and current of the blank are also shown. (b) Fraction modern of large, radiocarbon dead samples compared to sample current. Horizontal and vertical bars represent measurement uncertainty in current and ratio, respectively.", echo = FALSE}
knitr::include_graphics("../figures/fig3_blanks.pdf")
```

@@

@@

Reference materials run with the hybrid gas method were in good agreement with consensus values. Mean difference from consensus for all measurements was F^14^C `r format(cons_sum$fm_diff_mean, digits = 1, nsmall = 4)` shows that the hybrid method has no bias in aggregate. The standard deviation of the difference in F^14^C was `r format(cons_sum$fm_diff_sd, nsmall = 4)`, which may be taken as an estimate of the 1$\sigma$ precision of the method. The mean of propagated measurement errors for these samples was `r format(cons_sum$sig_fm_corr_mean, nsmall = 4)` F^14^C for these samples, which is in good agreement with the method precision above. The method error is about three times larger than the roughly 3 ‰ precision of graphite measurements, but is within the range of method precision of other gas sources.

```{r tab1, message = FALSE}
# Summary table
cons_df %>% 
  group_by(Name, fm_consensus) %>% 
  summarize(across(c(fm_corr, sig_fm_corr, fm_diff, sigma),
                   list(mean = mean, sd = sd)),
            N = n()) %>% 
  ungroup() %>% 
  arrange(fm_consensus) %>% 
  mutate (fm_corr_err = set_errors(fm_corr_mean, fm_corr_sd),
          Name = case_when(Name == "TIRI-I" ~ "TIRI I",
                           Name == "C-2" ~ "IAEA C2",
                           Name == "NOSAMS2" ~ "NOSAMS 2")) %>% 
  select(`Sample Name` = Name, 
         `Reference F^14^C` = fm_consensus,
         `HGIS F^14^C ± stdev` = fm_corr_err,
         `Mean per-sample error` = sig_fm_corr_mean,
         N) %>% 
  knitr::kable(digits = 3,
               caption = "\\label{tab1}Reference material precision and accuracy. Reference 
               values are community consensus values for TIRI I and C2, and the 
               NOSAMS internal graphite value for NOSAMS2. NOSAMS2 is a 
               homogenized bivalve molusk reference. The mean of per-sample 
               error estimates is included for comparison to the reproducibilty 
               of the samples.")

The difference between reference samples and graphite consensus does not vary systematically by F^14^C, indicating that the method has no significant bias with radiocarbon content (Figure 4). Each reference standard type agrees with graphite consensus values within the standard deviation of samples measured, with slightly more measurements agreeing with graphite within 1$\sigma$ uncertainties than would be expected statistically.

```{r fig4, fig.width = 5, fig.height = 5, fig.align='center', out.width="50%", fig.cap = "\\label{fig4}Agreement of standards with graphite consensus values. Difference of HGIS measurements from consensus for individual samples are plotted against their expected value. Expected values are spread to avoid overlapping points. Error bars represent one sigma sample uncertainty. The black line is the mean difference from expected value for all measurements, and the blue lines are one standard deviation of all measurements.", echo = FALSE}
knitr::include_graphics("../figures/fig4_consensus_samples.pdf")
```

@@

We found the HGIS and graphite measurements to be in good agreement, with no significant bias over the range in F^14^C of the core samples (0.4 to 1.15) (Figure 5a). Mean difference between hybrid and graphite measurements was -0.003 F^14^C and the standard deviation of this difference was 0.009 F^14^C. Mean per-sample 1$\sigma$ uncertainty was 0.010 F^14^C for HGIS measurements, compared with 0.002 F^14^C for graphite. Three samples were analysed by HGIS in duplicate. Two of the three pairs agreed within 1$\sigma$ and all three agreed within 2$\sigma$. There is an apparent, but not significant positive relationship between the HGIS-graphite difference and measured value. The higher graphite values at lower F^14^C may indicate an overcorrection of HGIS blank for these samples.

```{r fig5, fig.width = 5, fig.height = 5, fig.align='center', out.width="50%", fig.cap = "\\label{fig5}Comparison of Haiti core samples measured as graphite and via HGIS. (a) Difference between HGIS and graphite measurements of the same materials. Difference between HGIS and graphite measurements is shown on the vertical axis and plotted against the mean of the two measurements. Samples are offset on the horizontal axis for clarity. Error bars are measurement error for each sample. Graphite measurements are included to allow comparison of error estimates. The bottom row of plots show a comparison of age-depth models produced using HGIS (b) and graphite measurements (c).", echo = FALSE}
knitr::include_graphics("../figures/fig5_core_carbonates.pdf")
```

@@

# References {-}
